Location and Scale-Invariant Power Transformations for Transforming Data to

### 2024-09-16 manuscript/manuscript.Rmd
Added more details regarding datasets containing mostly numeric features that display a large shift.
Updated software references.

We used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
The $\lambda$ parameter for conventional power transformations (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location- and scale-invariant power transformations (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
The required algorithms were implemented in the `power.transform` R software package **ADD REFERENCE**.
Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.

@@

Robust location- and scale-invariant transformations are a suitable replacement for their conventional counterparts.
They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the distribution of a feature.
This is particularly relevant for automated data processing, where such issues may go unnoticed.
Compared to non-robust location- and scale-invariant transformations, real-world examples with outliers showed higher residual errors of transformed features. 
Robust transformations seek to minimise residual errors for the central part of the distribution, instead of the entire distribution, including outliers.
The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit of for -- nor detriment to -- model performance for location- and scale-invariant power transformations.
One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.

@@

Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE]. The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository. Data and results for the machine learning experiment are separately available from Zenodo [https://doi.org/10.5281/zenodo.13736671].

>>>>

We used simulated data to assess invariance to location and scale of the proposed power transformations, weighting for robust transformations, and to develop the empirical central normality test.
The $\lambda$ parameter for conventional power transformations (Eqn. \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}), as well as $\lambda$, $x_0$ and $s$ parameters for location- and scale-invariant power transformations (Eqn. \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) were estimated using the BOBYQA algorithm for derivative-free bound constraint optimisation [@Powell2009-zb] through maximum likelihood estimation.
The required algorithms were implemented in the `power.transform` R software package [@Zwanenburg2024-kq].
Of note, the `power.transform` package shifts feature values into the positive domain if negative or zero values are present for Box-Cox power transformations.

@@

Robust location- and scale-invariant transformations are a suitable replacement for their conventional counterparts.
They demonstrated robustness against outliers and prevent inaccurate transformations and potential numerical issues due to location and scale of the distribution of a feature.
This is particularly relevant for automated data processing, where such issues may go unnoticed.
Compared to non-robust location- and scale-invariant transformations, real-world examples with outliers showed higher residual errors of transformed features. 
Robust transformations seek to minimise residual errors for the central part of the distribution, instead of the entire distribution, including outliers.
The empirical central normality test indicated that robust transformations are better able to achieve central normality in the presence of outliers.
However, in a machine learning experiment of 232 real-world datasets that contained at least one numeric feature, we did not find a meaningful benefit of for -- nor detriment to -- model performance for location- and scale-invariant power transformations.
One reason may be that numeric features with large location shifts ($|\mu| > 1000.0$) were uncommon.
Of the 4886 numeric features in the 232 datasets, 266 (5%) features in 34 datasets had large location shifts, of which 200 appeared in just 2 datasets.
For the latter two datasets, the transformation method does not show significant difference between groups (Friedman test; $p > 0.05$).
```{r include = FALSE, eval = TRUE}

data <- .get_ml_experiment_data(manuscript_dir = manuscript_dir)
data <- data[dataset %in% c("Hill_Valley_with_noise", "Hill_Valley_without_noise")]

@@

Location- and scale-invariant power transformations were implemented in the `power.transform` package for R, which is available from [GitHub](https://github.com/oncoray/power.transform) and the [CRAN repository](https://cran.r-project.org/package=power.transform). The manuscript was created using R Markdown and is likewise available from the `power.transform` GitHub repository. Data and results for the machine learning experiment are separately available from [Zenodo](https://doi.org/10.5281/zenodo.13736671).
